A type-safe generalized editor calculus (Short Paper)

### 2023-07-02 Article.tex
Mere nyt

Although a cursor context accurately specifies the subtree where the cursor must reside, it does not inherently guarantee the presence of only one cursor within that subtree. Therefore, it becomes necessary to introduce the concept of well-formed abt's. We will do this informally, where a well-formed abt is simply an abt containing exactly one cursor. This is guarantied when an abt can be interpreted as $C[\dot{a}]$, meaning the context C, where $\ctxhole$ is substituted with a tree $\dot{a}$, where the cursor-operator is either the root $\dot{a}$, or one of the children of the root.

@@

\begin{example}\label{ex:well_formed_trees}
   We will now show the extension of \cref{fig:example_abstract_syntax}, where cursors are included, as shown in \cref{fig:example_well_formed_trees}. We note that $\hat{s}$ and $\hat{e}$ are trees without cursors.
    \federeboks{R C l}{
        \dot{s} & ::= & let $x = [\hat{e}]$ in $\hat{s}$ | let $x = \hat{e}$ in $[\hat{s}]$ | $[\hat{e}]$ | $[\hat{s}]$ \\
        \dot{e} & ::= & $[\hat{e}_1] + \hat{e}_2$ | $\hat{e}_1 + [\hat{e}_2]$ | $[\hat{e}]$
    }{\caption{Abstract syntax of abt's with exactly one cursor}\label{fig:example_well_formed_trees}}
    Given the syntax in \cref{fig:example_well_formed_trees} and the previously introduced cursor context, we can now determine the well-formedness of any abt in our language. For example the statement
    \begin{equation*}
        \text{let $x=[\hole{e}]$ in $x+5$}
    \end{equation*}
    is well-formed since it can be written as $C[\dot{e}]$ where
    \begin{align*}
        C &= \text{let $x=\ctxhole$ in $x+5$} \\
        \dot{e}&= [\hole{e}]
    \end{align*}
    It should be noted that there often are two valid interpretations of a tree $a$ as $C[\dot{a}]$. In this example, the other valid interpretation would be:
    \begin{align*}
        C' &= \ctxhole \\
        \dot{e}' &= \text{let $x=[\hole{e}]$ in $x+5$}
    \end{align*}
\end{example}

@@

Unlike the editor calculus by \cite{type_safe_structure_editor}, we do not utilize any form of structural congruence on editor expressions to define the semantics of trivial sequential composition or recursion. Instead we simply define these notions with their own transition rule (Seq-Trivial) and (Recursion)
\federeboks{c}{
    \infeshort{Cond-1}
            {a |= \phi}
            {\lr{\phi => E_1 | E_2, C[a]} \ltrans{\epsilon} \lr{E_1, C[a]}}
    \\\\
    \infeshort{Cond-2}
        {a \nvDash \phi}
        {\lr{\phi => E_1 | E_2, C[a]} \ltrans{\epsilon} \lr{E_2, C[a]}}
    \\\\
    \infeshort{Seq}
        {\lr{E_1, a} \ltrans{\alpha} \lr{E_1',a'}}
        {\lr{E_1 \ggg E_2, a} \ltrans{\alpha} \lr{E_1' \ggg E_2,a'}}
    \\\\
    \infeshort{Seq-Trivial}
        {}
        {\lr{\text{nil} \ggg E_2, a} \ltrans{\epsilon} \lr{E_2,a}}    
    \\\\
    \infeshort{Recursion}
        {}
        {\lr{\text{rec } x.E,a} \ltrans{\epsilon} \lr{E[x:=rec \ x.E], a}}
    \\\\
    \infeshort{Context}
        {a \ltrans{\pi} a'}
        {\lr{\pi.E, C[a]} \ltrans{\pi} \lr{E, C[a']}}
}{\caption{Reduction rules for editor expressions}\label{fig:editor_expression_rules}}
.

@@

\section{Encoding the generalized editor calculus in an extended lambda calculus} \label{sec:encoding}
In this section we attempt to soundly encode the generalized editor calculus using the simply typed $\lambda$-calculus (\stlc) as the base calculus. The calculus will be extended with various constructs such as pattern matching, pairs and the recursive \textit{fix} operator. To encode the abstract binding trees, we will only need the simply typed lambda calculus. To encode the contexts, we introduce pairs. To encode atomic prefix commands we further extend the calculus with pattern matching ($\lambda_{->,p}$). Lastly to encode modal logic and editor expressions we extend the calculus with a fixed point operator ($\lambda_{->,p,fix}$).

A concrete overview of the calculus and the constructs necessary to encode every central concept of the editor calculus can be seen in \cref{tab:encoding_requirements}.

@@

\subsection{Motivation}
The motivation behind encoding the generalized editor calculus in $\lambda_{->,p,fix}$ (or subsets thereof) is most importantly the type system it provides. If our encoding is sound, then any instance of the editor calculus will have a sound type system, regardless of the abstract syntax. This follows from the fact that the type system of $\lambda_{->,p,fix}$ is sound. Our editor calculus having a sound type system means that it will reject incorrect editor expressions.

Secondly a $\lambda_{->,p,fix}$ encoding provides a clear strategy for implementing any instantiation of the generalized editor calculus in a functional programming language.

\subsection{Abstract Binding Trees}
To encode abstract binding trees we only need to use the simply typed lambda-calculus. We extend the lambda-calculus with term constants $o$ for every $o \in \OP$ excluding cursors, and the base types $s$ for every $s \in \SORT$. 
Typing rules for operators can be seen on \cref{fig:abt_typing_rules},
where we can infer the type of an operator in the lambda-calculus by
its arity.  
\federeboks{C}{
    \infeshort{T-Operator}
        {o \in \OP \text{ and has arity } (\Vec{s_1}.s_1, \dots, \Vec{s_n}.s_n)s}
        {\Gamma |- o : (\Vec{s_1} -> s_1) -> \dots (\Vec{s_n} -> s_n) -> s}
}{
    \caption{Typing rules for operators.}
    \label{fig:abt_typing_rules}
}

@@

    \federeboks{C C}{
        \text{plus} : e \to e \to e
        &
        \text{num}[n] : e
        \\\\
        \text{var}[x] : e
        &
        \text{hole}_e : e
        \\\\
        \text{exp} : e \to s
        &
        \text{hole}_s : s
        \\\\
        \multicolumn{2}{C}{
            \text{let} : e \to (e \to s) \to s
        }
    }{
        \caption{Types for the encoded operators.}
        \label{fig:abt_example_type}
    }

@@

\subsection{Cursor Contexts}\label{sec:enc_cursor_contexts}

@@

\subsection{Atomic Prefix Commands}
To encode the atomic prefix commands we further extend the lambda calculus with pattern matching. As seen in \cref{fig:apc_syntactic_forms} we introduce the match construct where we match the term $M$ on the patterns $p$ resulting in the corresponding term $N$ if there is a match. As mentioned $p$ is the syntactic category describing the pattern we are trying to match. These patterns consist of variables, wildcards, operators, pairs, and lastly bindings which recursively match on the body of an abstraction. 
\federeboks{R C L l}{
    \multicolumn{4}{c}{Terms}
    \\
    M, N & ::= & \text{match}\ M\ \overrightarrow{p \to N} & (match construct)
    \\
    p & ::= & x              & (variable) \\
      & |   & \_             & (wildcard) \\
      & |   & o\ \Vec{p}     & (operator) \\
      & |   & (p_1, p_2)     & (pair) \\
      & |   & .p             & (binding)
}{\caption{Abstract syntax for representing atomic prefix commands, extending the lambda-calculus with a match construct.}\label{fig:apc_syntactic_forms}}
The reduction rules for the match construct can be seen in the full version of the paper.

We now define the auxiliary functions as seen in \cref{tab:zip_oper}. These function as the operations of a zipper data structure, which helps encode atomic prefix commands on abts. We also introduce the abbreviation $[a]$ to mean the cursor operator with the child $a$.

\federeboks{L}{
    $down$ \define \lambda x:s.\text{match }x
    \quad [o \ (.a_1) \dots (.a_n)] \to \\
    \quad o \ (\lambda \Vec{x_1}.[a_1]) \dots (\Vec{x_n}.a_n) \\\\

    $right$ \define \lambda x:s.\text{match }x
    \quad o \ (.a_1) \ \dots (.[a_i]) \ \dots (.a_n) \to \\
    \quad o \ (\lambda \Vec{x_1}.a_1) \ \dots (\lambda \Vec{x_{i+1}}.[a_{i+1}]) \ \dots (\Vec{x_n}.a_n) \\\\

    $up$ \define \lambda x:s.\text{match }x
    \quad o \ (.a_1) \dots (.[a_i]) \dots (.a_n) \to \\
    \quad [o \ (\lambda \Vec{x_1}.a_1) \dots (\lambda \Vec{x_n}.a_n)] \\\\

    $set$ \define \lambda a:s.\lambda x:s.\text{match }x
    \quad [a'] \to [a]
}{\caption{Function definitions for cursor movement and substitution on abstract binding trees.}\label{tab:zip_oper}}
The encoding of atomic prefix commands can be seen in \cref{fig:atc_enc}. The $child$ command can be encoded as a base case and a recursive case. In the base case we want to move to the first $child$, which means we just go down. In the recursive case we want to move to $child$ $n$, we can encode this as moving to $child$ $n-1$ and then moving right. $parent$ and $insert$ can directly be encoded as $up$ and $set$ respectively.
\federeboks{L L}{
    \encode{\text{child} \ 1} = \text{down}
    &
    \encode{\text{child} \ n} = \text{right}\ \encode{\text{child} \ n - 1}
    \\\\
    \encode{\text{parent}} = \text{up}
    &
    \encode{\text{insert} \ a} = \text{set} \ \encode{a}
}{
    \caption{Encoding of atomic prefix commands.}
    \label{fig:atc_enc}
}

\subsection{Control Structures and Modal Logic}
To encode control structures and modal logic, we need to further extend the lambda-calculus with a \fix operator as well as the $Bool$ base type. We introduce the \fix operator as seen in \cite{types_programming_languages}, as well as boolean types and values. 
On \cref{fig:edt_enc} the encoding of editor expressions can be seen. For atomic prefixes the atomic prefix command should be applied before the rest of the editor expression. The $nil$ editor expression will become the identity function. Sequential editor expressions will apply the encoding of the editor expressions in the correct order. Recursion works by using the newly introduced \fix operator. Lastly, conditional editor expressions are encoded such that we match on the evaluation of $\phi$ on the first projection of a context $C$ and apply the encoding of $E_1$ or $E_2$ on $C$ accordingly. For the full encoding of modal logic, we refer to the full version of the paper.
%\federeboks{R C L l}{
%    \multicolumn{4}{c}{Terms}
%    \\
%    M & ::= & \fix\ M & (fixed point of M) \\
%      & |   & \top    & (true) \\
%      & |   & \bot    & (false)
%    \\\\
%    p & ::= & \top & (match true) \\
%      & |   & \bot & (match false)
%    \\\\
%    \multicolumn{4}{c}{Types}
%    \\
%    \tau & ::= & \text{Bool} & (boolean)
%}{\caption{Abstract syntax for control structures and modal logic, extending the calculus with the \fix operator and boolean types and term constants.}\label{fig:control_syntactic_forms}}

@@

    \text{neg} &\define& \lambda b:\text{Bool} . \text{match } b \\
        & |     & \top \to \bot \\
        & |     & \bot \to \top
}{
    \caption{Function definitions for auxiliary functions.}
    \label{fig:cond_aux}
}
The encoding of conjunction, disjunction and negation, shown in \cref{fig:encoding_connectives}, comes quite naturally using the previously defined auxiliary functions.
\federeboks{L}{
    \encode{\phi_1 \wedge \phi_2 } = \text{phoenix and} \ \encode{\phi_1} \ \encode{\phi_2}
    \\\\

@@

    \encode{\neg \phi} = \text{bluebird neg} \ \encode{\phi}
    \\\\
    \encode{@o} = \lambda x:s . \text{match } x \\
                \quad | \ [o \ \_ \ \dots \ \_] \to \top  \\
                \quad | \ \_ \to \bot
}{\caption{Encoding of propositional connectives}\label{fig:encoding_connectives}}
In \cref{fig:cond_enc} we show the encoding of modal logic operators.
\federeboks{L}{    
    \encode{\Diamond o} = \text{\fix} (\lambda f:(s->\text{Bool})
    . \lambda x:s . \text{match } x \\
                        \quad | \ [o \ \_ \ \dots \ \_] \to \top \\
                        \quad | \  [\_(\_.a_1 , \dots ,\_.a_n)] \to \\
                        \quad\ \ \text{or} ( \dots (\text{or}\ (f\ [a_1]) \ (f \ [a_2]))\ \dots ) \ (f\ [a_n])\\
                        \quad | \ \_ \to \bot )
    \\\\
    \encode{\square o_1} = \lambda x:s . \text{match } x \\
                         \quad | \ \_(\_.a_1 , \dots ,\_.a_n) \to \\
                         \quad\ \ \text{and}(\dots (\text{and}  (\encode{\Diamond o}\ a_1) (\encode{\Diamond o} \ a_2)) \dots )\ (\encode{\Diamond o} \ a_n) \\
                         \quad | \ \_ \to \top
}{
    \caption{Encoding of modal logic and boolean logic operators.}
    \label{fig:cond_enc}
}
$@o$ is encoded such that we pattern match on the on the operator $o$, and if there is a match the resulting value is true, otherwise it is false. $\Diamond o$ uses the \fix operator, as the function is recursively defined. 
$\square o$ consists of either applying $\Diamond o$ on all the children of the tree, or returning true if there are no children.
\fi
With this, the entire editor calculus has been encoded. We refer to the full version of the paper to see the proof of the soundness of our encoding.
\iffalse
\begin{example}
    To illustrate how editor expressions are encoded, we provide an encoding of the following expression
    \begin{equation*}
        \text{child\ $2$.nil $\ggg$ parent.nil}
    \end{equation*}
    using the encoding rules seen on \cref{fig:edt_enc} and \cref{fig:atc_enc}. For readability purposes we will not add explicit typings, just note that any variable $C_i$ has type $Ctx$.
    \begin{align*}
         &\ \encode{\text{child\ $2$.nil $\ggg$ parent.nil}} \\
         =&\ \lambda C_1 . \encode{\text{parent.nil}} (\encode{\text{child\ $2$.nil}} C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . \encode{\text{nil}}(\encode{\text{parent}} C_2.1, C_2.2)) (\encode{\text{child\ $2$.nil}} C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . \encode{nil}(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (\lambda C_3 . C_3)(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
         \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . \encode{nil} (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\lambda C_4.C_4) (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
         \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (\encode{child\ 1} (C_3.1, C_3.2))C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (down (C_3.1, C_3.2))C_1) \\
    \end{align*}
    As can be seen, the editor expression could be fully encoded to the extended lambda calculus. We refer to the full paper for
\end{example}
\fi

\section{Conclusion \& Further Work}
%Conclusion
We have developed a generalized editor calculus that enables the creation of a syntax-directed editor calculus for a specific abstract syntax. Subsequently, we encoded this editor calculus into an extended version of the lambda calculus, which incorporates pairs, recursion and pattern matching.
Further work on our generalized editor calculus could be to prove whether the encoding is complete. Otherwise, one could implement a redo-undo construct such as introduced in \cite{undo_Edit} in the generalized editor calculus. Another interesting construct to implement would be a copy-paste operation.

@@

%We have identified several possibilities for expanding our work on the generalized editor calculus. First of all, it would be excellent to prove that our encoding is complete. However, in the scope of this paper, we decided to only include the proof of the soundness of the encoding, as that would give us the sound type system.

>>>>

Although a cursor context accurately specifies the subtree where the cursor must reside, it does not inherently guarantee the presence of only one cursor within that subtree. Therefore, it becomes necessary to introduce the concept of well-formed abt's. Informally, a well-formed abt is simply an abt containing exactly one cursor. This is guarantied when an abt can be interpreted as $C[\dot{a}]$, meaning the context C, where $\ctxhole$ is substituted with a tree $\dot{a}$, where the cursor-operator is either the root $\dot{a}$, or one of the children of the root.

@@

% \begin{example}\label{ex:well_formed_trees}
%    We will now show the extension of \cref{fig:example_abstract_syntax}, where cursors are included, as shown in \cref{fig:example_well_formed_trees}. We note that $\hat{s}$ and $\hat{e}$ are trees without cursors.
%     \federeboks{R C l}{
%         \dot{s} & ::= & let $x = [\hat{e}]$ in $\hat{s}$ | let $x = \hat{e}$ in $[\hat{s}]$ | $[\hat{e}]$ | $[\hat{s}]$ \\
%         \dot{e} & ::= & $[\hat{e}_1] + \hat{e}_2$ | $\hat{e}_1 + [\hat{e}_2]$ | $[\hat{e}]$
%     }{\caption{Abstract syntax of abt's with exactly one cursor}\label{fig:example_well_formed_trees}}
%     Given the syntax in \cref{fig:example_well_formed_trees} and the previously introduced cursor context, we can now determine the well-formedness of any abt in our language. For example the statement
%     \begin{equation*}
%         \text{let $x=[\hole{e}]$ in $x+5$}
%     \end{equation*}
%     is well-formed since it can be written as $C[\dot{e}]$ where
%     \begin{align*}
%         C &= \text{let $x=\ctxhole$ in $x+5$} \\
%         \dot{e}&= [\hole{e}]
%     \end{align*}
%     It should be noted that there often are two valid interpretations of a tree $a$ as $C[\dot{a}]$. In this example, the other valid interpretation would be:
%     \begin{align*}
%         C' &= \ctxhole \\
%         \dot{e}' &= \text{let $x=[\hole{e}]$ in $x+5$}
%     \end{align*}
% \end{example}

@@

% Unlike the editor calculus by \cite{type_safe_structure_editor}, we do
% not utilize any form of structural congruence on editor expressions to
% define the semantics of trivial sequential composition or
% recursion. Instead we simply define these notions with their own
% transition rule (Seq-Trivial) and (Recursion) 
\federeboks{c}{
    \infeshort{Cond-1}
            {a |= \phi}
            {\lr{\phi => E_1 | E_2, C[a]} \ltrans{\epsilon} \lr{E_1, C[a]}}
    \\\\
    \infeshort{Cond-2}
        {a \nvDash \phi}
        {\lr{\phi => E_1 | E_2, C[a]} \ltrans{\epsilon} \lr{E_2, C[a]}}
    \\\\
    \infeshort{Seq}
        {\lr{E_1, a} \ltrans{\alpha} \lr{E_1',a'}}
        {\lr{E_1 \ggg E_2, a} \ltrans{\alpha} \lr{E_1' \ggg E_2,a'}}
    \\\\
    \infeshort{Seq-Trivial}
        {}
        {\lr{\text{nil} \ggg E_2, a} \ltrans{\epsilon} \lr{E_2,a}}    
    \\\\
    \infeshort{Recursion}
        {}
        {\lr{\text{rec } x.E,a} \ltrans{\epsilon} \lr{E[x:=rec \ x.E], a}}
    \\\\
    \infeshort{Context}
        {a \ltrans{\pi} a'}
        {\lr{\pi.E, C[a]} \ltrans{\pi} \lr{E, C[a']}}
}{\caption{Reduction rules for editor expressions}\label{fig:editor_expression_rules}}
.

@@

\section{Encoding the generalized editor calculus} \label{sec:encoding}

To encode the abstract binding trees, the simply typed lambda calculus
suffices.

In order to encode the generalized editor calculus using the simply
typed $\lambda$-calculus (\stlc) as the base calculus, we extend the
latter with new constructs.

To encode the contexts, we introduce pairs. To encode atomic prefix
commands we further extend the calculus with pattern matching
($\lambda_{->,p}$). Finally, to encode modal logic and editor
expressions recursion is needed in the form of the fixed point
operator ($\lambda_{->,p,fix}$).

An overview of the calculus and the constructs necessary to encode
each part of it can be seen in \cref{tab:encoding_requirements}.

@@

 All of the lambda calculi that we consider in the following have a
 standard reduction semantics with reductions of the form $e \to
 e'$. Moreover, every calculus has a type systems with type judgements
 of the form $\Gamma \vdash e : \tau$, where $\tau$ ranges over a set
 of types \Types\ and $\Gamma$ is a type environment
 $\Gamma : \Var \pra \Types$. The type systems all satisfy the subject
 reduction property.

 \begin{theorem}[Subject reduction]\label{thm:subred}
   If $\Gamma \vdash e : \tau$ and $e \to e'$, then $\Gamma \vdash e'
   : \tau$.
 \end{theorem}

 It is this result that ensures that our encoding will indeed provide
 a general result about type soundness for the generalized editor
 calculus.

\subsection{Abstract binding trees}
To encode abstract binding trees we only need to use the simply typed lambda-calculus. We extend the lambda-calculus with term constants $o$ for every $o \in \OP$ excluding cursors, and the base types $s$ for every $s \in \SORT$. 
Typing rules for operators can be seen on \cref{fig:abt_typing_rules},
where we can infer the type of an operator in the lambda-calculus by
its arity.  
\federeboks{C}{
    \infeshort{T-Operator}
        {o \in \OP \text{ and has arity } (\Vec{s_1}.s_1, \dots, \Vec{s_n}.s_n)s}
        {\Gamma |- o : (\Vec{s_1} -> s_1) -> \dots (\Vec{s_n} -> s_n) -> s}
}{
    \caption{Typing rules for operators.}
    \label{fig:abt_typing_rules}
}

@@

    \federeboks{C C}{
        \text{plus} : e \to e \to e
        &
        \text{num}[n] : e
        \\\\
        \text{var}[x] : e
        &
        \text{hole}_e : e
        \\\\
        \text{exp} : e \to s
        &
        \text{hole}_s : s
        \\\\
        \multicolumn{2}{C}{
            \text{let} : e \to (e \to s) \to s
        }
    }{
        \caption{Types of the encoded operators.}
        \label{fig:abt_example_type}
    }

@@

\subsection{Cursor contexts}\label{sec:enc_cursor_contexts}

@@

\subsection{Atomic prefix commands}

To encode the atomic prefix commands we further extend the lambda
calculus with pattern matching. The formation rules are extended as
shown below; $p$ is the metavariable for patterns.
% As seen in
% \cref{fig:apc_syntactic_forms} we introduce the match construct where
% we match the term $M$ on the patterns $p$ resulting in the
% corresponding term $N$ if there is a match. As mentioned $p$ is the
% syntactic category describing the pattern we are trying to
% match. These patterns consist of variables, wildcards, operators,
% pairs, and lastly bindings which recursively match on the body of an
% abstraction.  
% \federeboks{R C L l}{
%     \multicolumn{4}{c}{Terms}
%     \\
\begin{align*}
    M, N & ::= & \text{match}\ M\ \overrightarrow{p \to N} && \text{(match construct)}
    \\
    p & ::= & x              & & \text{(variable)} \\
      & |   & \_             && \text{(wildcard)} \\
      & |   & o\ \Vec{p}     && \text{(operator)} \\
      & |   & (p_1, p_2)     && \text{(pair)} \\
         & |   & .p             && \text{(binding)}
\end{align*}
%} {\caption{Abstract syntax for representing atomic prefix commands, extending the lambda-calculus with a match construct.}\label{fig:apc_syntactic_forms}}
%The reduction rules for the match construct can be seen in the full version of the paper.
%
We are now able to define functions functions that describe navigation
in an abt and substitution; the definitions are givren in
\cref{tab:zip_oper}. We introduce the abbreviation $[a]$ to mean the
cursor operator with the child $a$.

\federeboks{L}{
    \textit{down} \define \lambda x:s.\text{match }x
    \quad [o \ (.a_1) \dots (.a_n)] \to \\
    \quad o \ (\lambda \Vec{x_1}.[a_1]) \dots (\Vec{x_n}.a_n) \\\\

    \textit{right} \define \lambda x:s.\text{match }x
    \quad o \ (.a_1) \ \dots (.[a_i]) \ \dots (.a_n) \to \\
    \quad o \ (\lambda \Vec{x_1}.a_1) \ \dots (\lambda \Vec{x_{i+1}}.[a_{i+1}]) \ \dots (\Vec{x_n}.a_n) \\\\

    \textit{up} \define \lambda x:s.\text{match }x
    \quad o \ (.a_1) \dots (.[a_i]) \dots (.a_n) \to \\
    \quad [o \ (\lambda \Vec{x_1}.a_1) \dots (\lambda \Vec{x_n}.a_n)] \\\\

    \textit{set} \define \lambda a:s.\lambda x:s.\text{match }x
    \quad [a'] \to [a]
}{\caption{Defining cursor movement and
    substitution.}\label{tab:zip_oper}}

The encoding of atomic prefix commands are given in
\cref{fig:atc_enc}. Note that the $child$ command is defined
recursively. \textit{parent} and \textit{insert} are simply encoded as \textit{up} and \textit{set} respectively.
\federeboks{L L}{
    \encode{\textit{child} \ 1} = \textit{down}
    &
    \encode{\textit{child} \ n} = \textit{right}\ \encode{\text{child} \ n - 1}
    \\\\
    \encode{\textit{parent}} = \textit{up}
    &
    \encode{\textit{insert} \ a} = \textit{set} \ \encode{a}
}{
    \caption{Encoding of atomic prefix commands.}
    \label{fig:atc_enc}
}

\subsection{Control structures and modal logic}

To encode control structures and modal logic, we need to further
extend the lambda-calculus with a \fix operator as well as the $Bool$
base type. We introduce the \fix operator as seen in
\cite{types_programming_languages}, as well as boolean types and
values.

\cref{fig:edt_enc} shows the encoding of editor expressions. For
atomic prefixes the atomic prefix command is applied before the
rest of the editor expression. The $nil$ editor expression is
the identity function. Sequential editor expressions apply the
encoding of the editor expressions in the correct order. Recursion
works by using the \fix operator. Lastly, conditional
editor expressions are encoded such that we match on the evaluation of
$\phi$ on the first projection of a context $C$ and apply the encoding
of $E_1$ or $E_2$ on $C$ accordingly. 
%\federeboks{R C L l}{
%    \multicolumn{4}{c}{Terms}
%    \\
%    M & ::= & \fix\ M & (fixed point of M) \\
%      & |   & \top    & (true) \\
%      & |   & \bot    & (false)
%    \\\\
%    p & ::= & \top & (match true) \\
%      & |   & \bot & (match false)
%    \\\\
%    \multicolumn{4}{c}{Types}
%    \\
%    \tau & ::= & \text{Bool} & (boolean)
%}{\caption{Abstract syntax for control structures and modal logic, extending the calculus with the \fix operator and boolean types and term constants.}\label{fig:control_syntactic_forms}}

@@

    \text{neg} &\define& \lambda b:\text{Bool} . \text{match } b \\
        & |     & \top \to \bot \\
        & |     & \bot \to \top
}{
    \caption{Function definitions for auxiliary functions.}
    \label{fig:cond_aux}
}
The encoding of propositional connectives is shown in
\cref{fig:encoding_connectives}.

\federeboks{L}{
    \encode{\phi_1 \wedge \phi_2 } = \text{phoenix and} \ \encode{\phi_1} \ \encode{\phi_2}
    \\\\

@@

    \encode{\neg \phi} = \text{bluebird neg} \ \encode{\phi}
    \\\\
    \encode{@o} = \lambda x:s . \text{match } x \\
                \quad | \ [o \ \_ \ \dots \ \_] \to \top  \\
                \quad | \ \_ \to \bot
}{\caption{Encoding of propositional connectives}\label{fig:encoding_connectives}}
In \cref{fig:cond_enc} we show the encoding of modal logic operators.
\federeboks{L}{    
    \encode{\Diamond o} = \text{\fix} (\lambda f:(s->\text{Bool})
    . \lambda x:s . \text{match } x \\
                        \quad | \ [o \ \_ \ \dots \ \_] \to \top \\
                        \quad | \  [\_(\_.a_1 , \dots ,\_.a_n)] \to \\
                        \quad\ \ \text{or} ( \dots (\text{or}\ (f\ [a_1]) \ (f \ [a_2]))\ \dots ) \ (f\ [a_n])\\
                        \quad | \ \_ \to \bot )
    \\\\
    \encode{\square o_1} = \lambda x:s . \text{match } x \\
                         \quad | \ \_(\_.a_1 , \dots ,\_.a_n) \to \\
                         \quad\ \ \text{and}(\dots (\text{and}  (\encode{\Diamond o}\ a_1) (\encode{\Diamond o} \ a_2)) \dots )\ (\encode{\Diamond o} \ a_n) \\
                         \quad | \ \_ \to \top
}{
    \caption{Encoding of modal logic and boolean logic operators.}
    \label{fig:cond_enc}
}
$@o$ is encoded such that we pattern match on the on the operator $o$, and if there is a match the resulting value is true, otherwise it is false. $\Diamond o$ uses the \fix operator, as the function is recursively defined. 
$\square o$ consists of either applying $\Diamond o$ on all the children of the tree, or returning true if there are no children.
\fi

% \begin{example}
%     We are now able to give an example of how to encode editor
%     expressions. Consider
%     \begin{equation*}
%         \text{child\ $2$.nil $\ggg$ parent.nil}
%     \end{equation*}
%     using the encoding rules seen on \cref{fig:edt_enc} and \cref{fig:atc_enc}. For readability purposes we will not add explicit typings, just note that any variable $C_i$ has type $Ctx$.
%     \begin{align*}
%          &\ \encode{\text{child\ $2$.nil $\ggg$ parent.nil}} \\
%          =&\ \lambda C_1 . \encode{\text{parent.nil}} (\encode{\text{child\ $2$.nil}} C_1) \\
%          =&\ \lambda C_1 . (\lambda C_2 . \encode{\text{nil}}(\encode{\text{parent}} C_2.1, C_2.2)) (\encode{\text{child\ $2$.nil}} C_1) \\
%          =&\ \lambda C_1 . (\lambda C_2 . \encode{nil}(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
%          =&\ \lambda C_1 . (\lambda C_2 . (\lambda C_3 . C_3)(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
%          \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
%          =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . \encode{nil} (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
%          =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\lambda C_4.C_4) (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
%          \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
%          =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (\encode{child\ 1} (C_3.1, C_3.2))C_1) \\
%          =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (down (C_3.1, C_3.2))C_1) \\
%     \end{align*}
%     As can be seen, the editor expression could be fully encoded to the extended lambda calculus. We refer to the full paper for
% \end{example}

\subsection{Soundness of the encoding}

The main result about our encoding is that it is sound wrt. the
operational semantics of the original calculus.

\begin{theorem}[Soundness] For every editor expression $E$ and abt $a$
  we have that if
  %
  \[ \lr{E,a} \ltrans{} \lr{E',a'} \]
  then
  \[ \encode{E}\encode{a} \rightarrow^{\ast}  \encode{E'}\encode{a'}  \]
\end{theorem}

\section{Conclusion and Further Work}

We have developed a generalized editor calculus that enables the
creation of a syntax-directed editor calculus for a specific abstract
syntax. Subsequently, we encoded this editor calculus into an extended
version of the lambda calculus, which incorporates pairs, recursion
and pattern matching.  Further work on our generalized editor calculus
could be to prove whether the encoding is complete. Otherwise, one
could implement a redo-undo construct such as introduced in
\cite{undo_Edit} in the generalized editor calculus. Another
interesting construct to implement would be a copy-paste operation.

@@

We have identified several possibilities for expanding our work on the
generalized editor calculus. First of all, the encoding should be
proved complete. Secondly,
